---
title: "EDA"
output: github_document
---
```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r loading dataset}
weather_df = read_csv(file = "./data/weather_df.csv") %>%
  mutate(
    month = lubridate::floor_date(date, unit = "month")
  )
weather_df
```

## 'group_by'

```{r}
weather_df %>%
  group_by(name, month)
```

## counting things

count month observations

```{r}
weather_df %>%
  group_by (name, month) %>%
  summarize(n_obs = n())

```

we can use 'count()'

```{r}
weather_df %>%
  count(name, month, name = "n_obs")
```

**NEVER** use base R's 'table'

the professor hates this code and we should not use it
```{r}
weather_df %>%
  pull(month) %>%
  table()
```


other helpful counters
```{r}
weather_df%>%
  group_by(month)%>%
  summarize(
    n_obs=n(),
    n_days=n_distinct(date)
  )
```

## a digression on 2x2 tables

```{r}
weather_df %>%
  filter(name != "Molokai_HI") %>%
  mutate(
    cold=case_when(
     tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  )%>%
  group_by(name, cold) %>%
  summarize(count = n())
```


```{r}
weather_df %>%
  filter(name != "Molokai_HI") %>%
  mutate(
    cold=case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  )%>%
  janitor::tabyl(name, cold)
```

## general summaries

the na.rm helps us drop blank ones 

you can do lots of summaries. 

```{r}
weather_df %>%
  group_by(name, month)%>%
  summarize(
    mean_tmax = mean(tmax, na.rm=TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, no.rm=TRUE)
  )
        
```

this is a dataframe!
```{r}
weather_df %>%
  group_by(name, month)%>%
  summarize(
    mean_tmax = mean(tmax, na.rm=TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, no.rm=TRUE)
  ) %>%
  ggplot(aes(x=month, y=mean_tmax, color=name)) +
           geom_point()+
            geom_line()
```

suppose you want to summarize many columns
```{r}
weather_df%>%
  group_by(name, month)%>%
  summarize(across(prcp:tmin, mean))
```

reminder: sometimes your results are easier to read in another format

```{r}
weather_df %>%
  group_by(name, month)%>%
  summarize(mean_tmax = mean(tmax)) %>%
  pivot_wider(
    names_from=name,
    values_from=mean_tmax
  )%>%
knitr::kable(digits=1)
```

