---
title: "Visualization"
subtitle: "Part 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

```{r loading dataset}
weather_df = read_csv(file = "./data/weather_df.csv")
  
weather_df
```

## Remember this plot ...?

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)
```

## Labels

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )
```

## Scales

start with the same plot

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  )+
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )

```

Look at color scales

first example
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 scale_color_hue(h=c(100,300))

```

changing the title of key using name argument
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 scale_color_hue(
   name = "location",
   h=c(100,300))

```

professor's preferred method
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)

```

## Themes

shift the legend 

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
  theme(legend.position = "bottom")
  

```

change the overall theme

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
  theme_bw()
  
```

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
  theme_minimal()
  
```

other ones are theme_classic() that has no gridlines, a bunch in ggthemes package (theme_economist, theme_excel, etc.)

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
 ggthemes::theme_excel()
  
```

professor likes theme_minimal best

applying theme after other things will override those original things. so you need to tweak the overall theme. for example with moving the legend

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
theme(legend.position = "bottom")
  
```

## Setting options

this is something/options he would normally put at the very beginning of a document
```{r, eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal()+theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

