---
title: "Visualization"
subtitle: "Part 1"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

```{r loading dataset}
weather_df = read_csv(file = "./data/weather_df.csv")
  
weather_df
```

## Scatterplots!!

Create my first scatterplot ever.

```{r}
ggplot(weather_df, aes(x=tmin, y=tmax)) +
  geom_point()
```

new approach, same plot.

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point()
```

save and edit a plot object.

```{r}
weather_plot =
    weather_df %>%
   ggplot(aes(x=tmin, y=tmax))

weather_plot + geom_point()
```

## Advanced scatterplot ...

Start with the same one and make it fancy!

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point()+
  geom_smooth(se = FALSE)
```

What about the 'aes' placement?

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

Let's facet some things!!

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, alpha=tmin, color = name)) +
  geom_point()+
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(. ~ name)
```
** you can put alpha global option within geom_point or you can set it equal to a variable in the ggplot statement

Let's combine some elements and try a new plot.

```{r}
weather_df %>%
    ggplot(aes(x=date, y=tmax, color=name))+
  geom_point(aes(size=prcp),alpha=.5)+
  geom_smooth(se=FALSE)+
  facet_grid(.~name)
```
## Some small notes

How many geoms have to exist?

You can have whatever geoms you want

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_smooth(se=FALSE)
```

You can use a neat geom!

```{r}
weather_df %>%
  ggplot (aes(x=tmin, y=tmax))+
  geom_hex()
```

```{r}
weather_df %>%
  ggplot (aes(x=tmin, y=tmax))+
  geom_bin2d()
```
```{r}
weather_df %>%
  ggplot (aes(x=tmin, y=tmax))+
  geom_density2d()+
  geom_point(alpha=.3)
```

## Univariate plots

Histograms are really great.

```{r}
weather_df %>%
  ggplot(aes(x=tmin))+
  geom_histogram()
```

Can we add color....

```{r}
weather_df %>%
  ggplot(aes(x=tmin, fill=name))+
  geom_histogram()+
  facet_grid(.~name)
```
```{r}
weather_df %>%
  ggplot(aes(x=tmin))+
  geom_histogram()+
  facet_grid(.~name)
```

```{r}
weather_df %>%
  ggplot(aes(x=tmin, fill=name))+
  geom_histogram(position = "dodge")
```

Let's try a new geometry!

```{r}
weather_df %>%
  ggplot(aes(x=tmin, fill=name))+
  geom_density(alpha=.4)
```

What about box plots?

```{r}
weather_df %>%
  ggplot(aes(x=name, y=tmin))+
  geom_boxplot()
```

trendy plots...


```{r}
weather_df %>%
  ggplot(aes(x=name, y=tmin, fill=name))+
  geom_violin()+
  stat_summary()
```

for stat_summary, it will default to the mean but you can also do it for the median

ridge plots --- the most popular plot of 2017

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=name))+
  geom_density_ridges()
```

## Save and embed

let's save a scatterplot.

```{r}
weather_plot =
  weather_df %>%
  ggplot(aes(x=tmin, y=tmax,color=name))+
  geom_point(alpha=.5)

ggsave("weather_plot.pdf", weather_plot, width=5, height=8)
```

What about embeding...
```{r, fig.width=6, fig.height=3}
weather_plot
```

embed at different size
```{r, fig.width=12, fig.height=6}
weather_plot
```



