---
title: "Visualization"
subtitle: "Part 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

```{r loading dataset}
weather_df = read_csv(file = "./data/weather_df.csv")
  
weather_df
```

## Remember this plot ...?

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)
```

## Labels

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )
```

## Scales

start with the same plot

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  )+
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )

```

Look at color scales

first example
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 scale_color_hue(h=c(100,300))

```

changing the title of key using name argument
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 scale_color_hue(
   name = "location",
   h=c(100,300))

```

professor's preferred method
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)

```

## Themes

shift the legend 

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
  theme(legend.position = "bottom")
  

```

change the overall theme

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
  theme_bw()
  
```

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
  theme_minimal()
  
```

other ones are theme_classic() that has no gridlines, a bunch in ggthemes package (theme_economist, theme_excel, etc.)

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
 ggthemes::theme_excel()
  
```

professor likes theme_minimal best

applying theme after other things will override those original things. so you need to tweak the overall theme. for example with moving the legend

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha=.5)+
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (C)",
    y = "Max daily temp (C)",
    caption = "Data from rnoaa package; temperatures in 2021"
  )+
 viridis::scale_color_viridis(
   name = "location",
   discrete=TRUE)+
theme(legend.position = "bottom")
  
```

## Setting options

this is something/options he would normally put at the very beginning of a document
```{r, eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal()+theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in 'geom'

basically below we creat two data sets, then apply two geometries. in the second example, we apply each geometry to a different dataset

```{r}
central_park =
  weather_df %>%
  filter(name == "CentralPark_NY")

molokai =
  weather_df %>%
  filter(name == "Molokai_HI")

ggplot(data = molokai, aes(x=date, y=tmax, color=name))+
  geom_point()+
  geom_line()
```

```{r}

ggplot(data = molokai, aes(x=date, y=tmax, color=name))+
  geom_point()+
  geom_line(data = central_park)
```

## Patchwork

remeber faceting? it lets you put the same plot types side by side

```{r}
weather_df %>%
  ggplot(aes(x=tmin, fill=name))+
  geom_density(alpha=.5)+
  facet_grid(.~name)
```

what happens when yuou want multipanel plots but can't facet .....?

```{r}
tmin_tmax_p=
  weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=.5)+
  theme(legend.position = "none")

prcp_dens_p =
  weather_df %>%
  filter(prcp > 0) %>%
  ggplot(aes(x=prcp, fill=name))+
  geom_density(alpha=.5)

tmax_date_p =
  weather_df %>%
  ggplot(aes(x=date, y=tmax, color=name))+
  geom_point()+
  geom_smooth(se=FALSE)+
  theme(legend.positon = "none")

tmin_tmax_p + prcp_dens_p + tmax_date_p
```

can also adjust how they show up in the panels in various ways

```{r}
tmin_tmax_p /( prcp_dens_p + tmax_date_p)
```

## Data Manipulation

Control your factors.

```{r}
weather_df %>%
  mutate(
    name = factor(name),
    name = forcats:: fct_relevel(name, c("Molokai_HI"))
  ) %>%
  ggplot(aes(y=tmax, x=name, fill=name))+
  geom_violin(alpha=.5)
```

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>%
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
  ) %>%
  ggplot(aes(x=temperature, fill = observation))+
  geom_density(alpha=.5)+
  facet_grid(. ~ name)
```




